---
title: "Applying more dplyr to my data”"
author: "Sierra Hess"
output: html_document
---

## Introduction 
In this code, I will tidy up my data by editing column names and will use functions that allow me to use uniform units and formatting in my data. 

## Set things up:
Add code to clear R's brain and to turn on the dplyr and here packages
```{r}
rm(list = ls()) # clear R's brain
library(dplyr)
library(ggplot2)
library (here)
```

## Read in data
Import the csv file called "day_one_data_for_homework13.csv" using stringAsFactors = True. 
Then use View(DF) to open the data in a new tab. 
```{r}
DF <- read.csv(here("Data", "day_one_data_for_homework13.csv"), stringsAsFactors = TRUE)
View(DF)
```

## Scroll through the data. Notice what happened to the column names that have spaces and the column that had a # in the column name in the .csv.  How did R change them?
R changed the column names with spaces and special characters by adding a point wherever an unreadable character was.

### Use the names() function (hint - chapter 2) to change the column names to, in order from left to right: Name, Date, Part, Tree_num, Species, leaf_length, leaf_width, leaf_units
```{r}
names(DF) <- c( "Name", "Date", "Part", "Tree_num", "Species", "leaf_length", "leaf_width", "leaf_units")
names(DF)
```

## Notice that some of the leaf measurements are in inches, and some are in metric units. Write code that selects just the rows of data that are in English units (inches) and that saves them to a new data frame called "English"

```{r}
English <-  filter(DF, leaf_units == "inches")
```
## Now notice that the dates have been entered in several formats, and that for one person, the years associated with the dates are incorrect. Write code that collects all of the data for which the dates have been formatted in the same way together into a single data frame and call it "Uniform_dates"

```{r}
Uniform_dates <- filter(DF, Date == "8/24/22")
```





