---
title: "Homework 11"
author: "Sierra Hess"
output: html_document
---

In this homework, I will learn how to use several new functions.
First, I will clear R's brain and download the libraries that I will use.
Set things up:
```{r}
rm(list = ls()) # clear R's brain
library(dplyr)
library(ggplot2)
library (here)
```

Next, I will read in my data using the here function.  and will use glimpse to get information about my data. 
```{r}
compensation <- read.csv(here("Data", "compensation.csv"))
glimpse(compensation)
```
Next, I will use the summary function in order to see the statistics of my data.
```{r}
summary(compensation)
```
Next, I will use the name function to get the names of the columns and the select function to grab columns.
```{r}
names(compensation)
select(compensation, Fruit) # use the fruit column
```
Next, I will select all columns except for one.
```{r}
select(compensation, -Root) #that is a minus sign
``` 

Now, I will use the slice function to grab rows. 
```{r}
slice(compensation, 2)
```

Next, I will invoke : in order to generate the sequence from the second to the tenth row. 
```{r}
slice(compensation, 2:10)
```

Now, I will collect row numbers with the c() function
```{r}
slice(compensation, c(2, 3, 10))
```
Next, I will use with to see how R interprets >.
```{r}
with(compensation, Fruit > 80)
```
Next, I will use filter() in order to return the true data for Fruit > 80
```{r}
filter(compensation, Fruit > 80)
```
Next, I will use an or function in order to add a second condition. 
```{r}
filter(compensation, Fruit > 80 | Fruit < 20)
```

Next, I will save my filtered data to an object. 
```{r}
lo_hi_fruit <- filter(compensation, Fruit > 80 | Fruit < 20)
lo_hi_fruit
```

Now I will use the head function to limit the rows that we see. 
```{r}
head(compensation)
```
Next, I will use mutate to designate a new column name and formation to my data.
```{r}
compensation <- mutate(compensation, logFruit = log(Fruit))
```
Next, I will use the head function again on the mutated data 
```{r}
head(compensation)
```
Next, I will use the arrange function to reorder the data into increasing fruit production. 
```{r}
arrange(compensation, Fruit)
```

Next, I will use the filter and select functions on the same line of code. 
```{r}
select(filter(compensation, Fruit > 80), Root)
```

Next I will use a pipe command in order to put mu commands in different rows. 
```{r}
compensation %>%
  filter(Fruit > 80) %>%
    select (Root)
```

Next, I will use the summarize and group_by functions in order to use the nested approach with no pipe. 
I will use the mean function to get the mean if the fruit. 
```{r}
summarize(
  group_by(compensation, Grazing),
        meanFruit = mean(Fruit))
```
Now I will make mean fruit an object
```{r}
mean.fruit <- summarize(
  group_by(compensation, Grazing),
        meanFruit = mean(Fruit))
```

Next, I will use the summarize function inside of a list rather than outside. 

```{r}
compensation %>%
  group_by(Grazing) %>%
  summarise(meanFruit = mean(Fruit))
```
Next, I will ask for more than one statistic. 

```{r}
compensation %>%
  group_by(Grazing) %>%
  summarise(
    meanFruit = mean(Fruit),
    sdFruit = sd(Fruit))
```