---
title: "HW 26"
author: "Sierra Hess"
date: "11/2/2022"
output: html_document
---

Set things up 
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Now lets pull in the data 
```{r}
soil <- read.csv(here("Data", "soil_reduced-1.csv"), stringsAsFactors = TRUE)
```
Before we begin any analysis, we need to check our data for possible errors/ outliers. Lets begin with that. 

A. Forest names
```{r}
levels(soil$Forest)
```

B. Lets check soil organic matter and soil calcium
start with a histogram
```{r}
ggplot(soil, aes(x = SOM_percent)) + 
  geom_histogram()
```

Now check soil calcium
```{r}
ggplot(soil, aes(x = Ca_ppm)) + 
  geom_histogram()
```

Lets run a summary on all of the variables:
```{r}
summary(soil)
```


## We'd like to ask the following questions from the data:

### 1. Is there a relationship between Soil organic matter (SOM, %) and soil calcium (Ca_ppm)?

Both variables are continuous 

Lets make a scatter plot

#### Plot relationship
```{r}
ggplot(soil, aes(x = SOM_percent, y = Ca_ppm, color = Forest)) + 
  geom_point() +
  theme_bw()
```

From looking at the data, it appears that there may be a weak positive relationship between SOM and Ca

Can I guess the SOM intercept?
Slope c. 1000/53 = 18ish , y int c. -5? 0? right around there. 

What statistical model fits these data. 
Both the reponse and predictor are continuous, so linear regression seems appropriate. 

So, lets construct the model:
```{r}
SOM_Ca_Mod <- lm(Ca_ppm ~ SOM_percent, soil)
summary(SOM_Ca_Mod)
```
Now I need to check the model assumptions before interpreting. 
```{r}
autoplot(SOM_Ca_Mod)
```

Based on my observations of the autoplot results, I'm modestly comfortable that my data meet the lm assumptions. 

Now look at model outcome. 
```{r}
anova(SOM_Ca_Mod)
```
```{r}
summary(SOM_Ca_Mod)
```
The p value is  > .05, so we cannot reject the null hypothesis which means that there is no clear relationship between soil organic matter and soil calcium. 


### 2. Does the abundance of worms differ between Donnerville and the Kip Tract?
Lets first look at a scatterplot of the data
```{r}
ggplot(soil, aes(x = Worms, y = Forest)) + 
  geom_point() +
  theme_bw()
```

Here, we can see that there does appear to be an average higher number of worms on the kip trail than on the  Donnerville path. 
Now, lets look at the mean number of worms on each trail. 
```{r}
sum1 <- soil %>%
  group_by(Forest) %>%
  summarise(meanWorms = mean(Worms))
sum1
```

Here, we can see that the Donnerville worms are much lower than the kip trail worms, so the abundance does differ. 



### 3. Does the abundance of soil Magnesium (Mg_ppm) differ between Peavine and South Hammond?



```{r}
sum2 <- soil %>%
  group_by(Forest) %>%
  summarise(meanMg = mean(Mg_ppm))
sum2
```

Here we can see that there is also a fairly large difference between the Peavine and south Hammond average magnesium.

### 4. Does the abundance of zinc (Zn_ppm) differ among forests?

Lets ignore na in kip trail Zn
```{r}
soil <- soil %>%
  filter(Zn_ppm >= 0)
```

Lets look at the average zinc abundance in each forest
```{r}
sum2 <- soil %>%
  group_by(Forest) %>%
  summarise(meanZn = mean(Zn_ppm))
sum2
```
Here, we can see that the average zinc differed between forests. 


