---
title: "Homework 24 draft"
author: "Sierra Hess"
output: html_document
---

# My first chi-square analysis

# Clear the decks
```{r}
rm(list = ls())
```

#libraries to use
```{r}
library(dplyr)
library(ggplot2)
library (here)
library(Hmisc)
```

```{r}
ozone <- read.csv(here("Data", "ozone.csv"), stringsAsFactors = TRUE)
lady <- read.csv(here("Data", "ladybirds.csv"), stringsAsFactors = TRUE)
```
# check it out 
```{r}
glimpse(lady)
```

# use groupby and summarise to summarise the data
```{r}
totals <- lady %>%
  group_by(Habitat, colour) %>% 
  summarise(total.number = sum(number))
```
```{r}
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge')

```

## change the colors of the histogram

```{r}
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c(black = "black", red = "red"))
```

# use xtabs for the X^2 test
```{r}
lady.mat <- xtabs(number ~ Habitat + colour, data = lady)

lady.mat
```

# Use Chisq.test
```{r}
chisq.test(lady.mat)
```

This tells us that there is a very small probability that the pattern is by chance. 

# look at the names from chisq test
```{r}
lady.chi <- chisq.test(lady.mat)
names(lady.chi)
```

# look at expected values 
```{r}
lady.chi$expected
```

# look at ozone
```{r}
glimpse(ozone)
```

# evaluate assumotions of a two sample t test
```{r}
ggplot(ozone, aes (x = Ozone)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~Garden.location, ncol = 1) + 
  theme_bw()
```


# Do a t test
```{r}
t.test(Ozone ~ Garden.location, data = ozone)
```

# use the var test
```{r}
var.test(Ozone ~ Garden.location, data = ozone)
```