---
title: "Homework 25"
author: "Sierra Hess"
output: html_document
---

# My first chi-square analysis

# Clear the decks
```{r}
rm(list = ls())
```

#libraries to use
```{r}
library(dplyr)
library(ggplot2)
library (here)
library(Hmisc)
library(ggfortify)
```

# load the data
```{r}
plant_gr <- read.csv(here("Data", "plant.growth.rate.csv"), stringsAsFactors = TRUE)
daphnia <- plant_gr <- read.csv(here("Data", "Daphniagrowth.csv"), stringsAsFactors = TRUE)
```

# look at the data
```{r}
glimpse(plant_gr)
```
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() + 
  ylab("Plant Growth Rate (mm/ week)") + 
  theme_bw
```
I get an error for using theme_bw, so I can remove this from the code. 

```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() + 
  ylab("Plant Growth Rate (mm/ week)") 
```

# Use the lm function to fit the model
```{r}
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, 
                data = plant_gr)
```

# Use autoplot
```{r}
autoplot(model_pgr, smooth.colour = NA)
```

# Use anova to view the variance table
```{r}
anova(model_pgr)
```
# Use the summary function 
```{r}
summary(model_pgr)
```

# Make a plot with a regression line
```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, y = plant.growth.rate)) + 
  geom_point() + 
  geom_smooth(method = 'lm') + 
  ylab("Plant Growth Rate (mm/week)") + 
  theme_bw()
  
```


# Use glimpse to view the daphnia data
```{r}
glimpse(daphnia)
```
# plot daphnia data using a boxplot
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) + 
  geom_boxplot() 
```


# use coord_flip to flip axis

```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) + 
  geom_boxplot() + 
  coord_flip()
```

# Fit the model 
```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

# Use autoplot
```{r}
autoplot(model_grow, smooth.colour = NA)
```

# View the anova table
```{r}
anova(model_grow)
```

# View the summary
```{r}
summary(model_grow)
```
# get the mean growth rates
```{r}
sumDat <- daphnia %>%
  group_by(parasite) %>%
  summarise(meanGR = mean(growth.rate))
sumDat
```

# install packages 
```{r}
install.packages("devtools")
library(devtools)
install_github("sinhrks/ggfortify")
```







