---
title: "Homework 23"
author: "Sierra Hess"
date: "10/24/2022"
output: html_document
---

## Set things up:
Add code to clear R's brain and to turn on the dplyr and here packages
```{r}
rm(list = ls()) # clear R's brain
library(dplyr)
library(ggplot2)
library (here)
library(Hmisc)
```

```{r}
soil <- read.csv(here("Data", "soil_data.csv"), stringsAsFactors = TRUE)
```

# make a scatterplot for SOM percent
```{r}
ggplot(soil, aes(y = SOM_percent, x = Field.Name)) + 
  geom_point(size = 5) +
  ylab("SOM percent") +
  xlab("Field Name")
  theme_bw()
```

# create boxplot for SOM percent
```{r}
ggplot(soil, aes(y = SOM_percent, x = Field.Name)) + 
  geom_boxplot() +
  geom_point(size = 4, colour = 'lightgrey', alpha = 0.5) +
  xlab("Field Name") +
  ylab("SOM percent")
  theme_bw()
```


# create a histogram for SOM percent
```{r}
ggplot(soil, aes(x = SOM_percent)) + 
  geom_histogram(bins = 10)

```

# make a scatterplot for Ca ppm
```{r}
ggplot(soil, aes(y = Ca.ppm, x = Field.Name)) + 
  geom_point(size = 5) +
  ylab("Ca ppm") +
  xlab("Field Name")
  theme_bw()
```



# create boxplot for SOM percent
```{r}
ggplot(soil, aes(y = Ca.ppm, x = Field.Name)) + 
  geom_boxplot() +
  geom_point(size = 4, colour = 'lightgrey', alpha = 0.5) +
  ylab("Ca.ppm") +
  xlab("Field Name")
  theme_bw()
```


# create a histogram for Ca ppm
```{r}
ggplot(soil, aes(x = Ca.ppmusethis::create_github_token())) + 
  geom_histogram(bins = 10)

```
SOMETHING HAPPENED ON THAT GGPLOT - NOT SURE WHY THE USETHIS CODE IS IN IT.

ALSO, YOU WERE SUPPOSSED TO MAKE A NEW FOREST COLUMN BEFORE YOU STARTED ALL THE PLOTS. SEE THE KEY FOR HOW TO DO THAT.  ALSO SEE BELOW AS POSSIBLE CODE:

```{r}
#create new column
soil$Forest <-"" #creates new empty character column
#sort forests and fill in column
Degrasse<-"DEGRASSE*" #creates wildcard to search
fix <- grep(Degrasse, soil$field_name)
soil$Forest[fix]<- "Degrasse"

Don <-"DONNERVILLE*"
fix <- grep(Don, soil$field_name)
soil$Forest[fix]<- "Donnerville"

Kip <-"KIP*"
fix <-grep(Kip, soil$field_name)
soil$Forest[fix] <- "Kip Tract"

Pea <-"PEAVINE*"
fix <-grep(Pea, soil$field_name)
soil$Forest[fix] <- "Peavine"

SH <-"SOUTH HAMMOND*"
fix <-grep(SH, soil$field_name)
soil$Forest[fix] <- "South Hammond"

soil$Forest
```
















